Binomial of sum of Gaussians versus sum of binomial of Gaussians

### 2022-10-07 src/0_check_bin-gaussian/bin-gaussian.Rmd
Iterate report

---
title: "Binomial of sum of Gaussians versus sum of binomial of Gaussians"
author: "Adam Howes"
output: html_document
---

```{r}
#' Binomial of sum of Gaussians versus sum of binomial of Gaussians
#' IID latent field
#' Without aggregation:
#' `y = \sum_{i = 1}^n y_i$, $y_i \sim \text{Bin}(m, \rho_i)$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)`
#' With aggregation:
#' `y \sim \text{Bin}(nm, \bar \rho)$, $\bar \rho = \frac{1}{n} \sum_{i = 1}^n \rho_i$, $\rho_i \sim \text{Logitnormal}(\mu, \sigma^2)`

experiment <- function(nsim, n, phi_mean, phi_sd, y_sample_size) {
  phi <- rnorm(n, phi_mean, phi_sd)
  rho <- plogis(phi)
  m <- rep(y_sample_size, n)

@@

nsim <- 500
n <- 36

#' Low prevalence setting (disease)
result <- experiment(nsim, n, -2.5, 0.5, 30)
plot(result$rho) #' This is what the underlying latent field looks like for these settings

@@

sd(result$y)
sd(result$y_agg)
sd(result$y_agg) / sd(result$y) #' Ratio

#' Probability approximately 1/2 setting (elections)
#' As before but with different `phi_mean` settings:
result_half <- experiment(nsim, n, 0, 0.5, 30)
plot(result_half$rho)

@@

sd(result_half$y)
sd(result_half$y_agg)
sd(result_half$y_agg) / sd(result_half$y) #' Ratio

#' Conclusions, things to add:
#' * Standard deviation of aggregation tends to overestimate true standard deviation,
#' * Effect is larger in the small `\rho` setting than the large `\rho` setting
#' * Should test latent field with spatial structure e.g. ICAR latent field and so on
#' * What is the effect of varying `sd_phi`? What is the effect of varying `y_sample_size`?
```

>>>>

---
title: "Binomial of sum of Gaussians versus sum of binomial of Gaussians"
author:
- name: Adam Howes
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    df_print: paged
    code_folding: show
    theme: lumen
abstract: |
  **Background**

  **Task**
---

## Set-up

Consider an IID latent field.
Without aggregation:
$$
\begin{align*}
y &= \sum_{i = 1}^n y_i, \\
y_i &\sim \text{Bin}(m, \rho_i), \\
\rho_i &\sim \text{Logitnormal}(\mu, \sigma^2)
\end{align*}
$$
With aggregation:
$$
\begin{align*}
y &\sim \text{Bin}(nm, \bar \rho), \\
\bar \rho &= \frac{1}{n} \sum_{i = 1}^n \rho_i, \\
\rho_i &\sim \text{Logitnormal}(\mu, \sigma^2)
\end{align*}
$$

```{r}
experiment <- function(nsim, n, phi_mean, phi_sd, y_sample_size) {
  phi <- rnorm(n, phi_mean, phi_sd)
  rho <- plogis(phi)
  m <- rep(y_sample_size, n)

@@

nsim <- 500
n <- 36
```

## Low prevalence setting (disease)

```{r}
result <- experiment(nsim, n, -2.5, 0.5, 30)
plot(result$rho) #' This is what the underlying latent field looks like for these settings

@@

sd(result$y)
sd(result$y_agg)
sd(result$y_agg) / sd(result$y) #' Ratio
```

## Probability approximately 1/2 setting (elections)

As before but with different `phi_mean` settings:

```{r}
result_half <- experiment(nsim, n, 0, 0.5, 30)
plot(result_half$rho)

@@

sd(result_half$y)
sd(result_half$y_agg)
sd(result_half$y_agg) / sd(result_half$y) #' Ratio
```

## Conclusions, things to add

* Standard deviation of aggregation tends to overestimate true standard deviation,
* Effect is larger in the small `rho` setting than the large `rho` setting
* Should test latent field with spatial structure e.g. ICAR latent field and so on
* What is the effect of varying `sd_phi`? What is the effect of varying `y_sample_size`?
